---
title: "Reproducible Research: Peer Assessment 1"
author: "julieb"
git_repo: "https://github.com/julieb-dreamlab/RepData_PeerAssessment1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r load, echo=TRUE}
library(dplyr)
library(ggplot2)
actdata <- read.csv("../data/activity.csv")

```


## What is mean total number of steps taken per day?
The total number of steps each day is calculated.
Report the summary of the daily totals

```{r sums, echo=TRUE}
#The total number of steps each day is calculated.
sm0<- with(actdata, tapply(steps,date,sum, na.rm=TRUE))
sumsdf<-data.frame(sm0)
names(sumsdf)<- c("sums")
c<-ggplot(sumsdf, aes(sums))
d<- c+ geom_histogram(binwidth = 1000)
d<- d+ labs(x= "Total Steps in a Day")
d<- d+ labs(y= "Count of Days")
d<- d+ labs(title= "Histogram of Total Steps in a Day")
print(d)
# report the summary of the daily totals
summary(sumsdf)
```

## What is the average daily activity pattern?
Sum the number of steps at each 5-minute interval across all days to see high activity periods.
```{r intervals, echo=TRUE}
#The steps per 5-minute interval are calculated.
sm1<- with(actdata, tapply(steps,interval,sum, na.rm=TRUE))
intsdf<-data.frame(sm1)
names(intsdf)<- c("sums")
intsdf<- mutate(intsdf,interval = as.numeric(rownames(intsdf)))
c<-ggplot(intsdf, aes(interval,sums))
d<- c+ geom_line()
d<- d+ labs(x= "Time of Day, minutes")
d<- d+ labs(y= "Total Steps during an Interval")
d<- d+ labs(title= "Daily Activity")
print(d)
maxint<-with(intsdf, intsdf[which(sums == max(sums)),])
```
The 5-minute interval where the maximum average number of step taken is {r maxint} (which starts at ).

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
